---
title: "BCM Datathon 2022"
author: "Andrew Zimolzak"
date: "November, 2022"
output: pdf_document
---


```{r initialize, message=FALSE, warning=FALSE, include=FALSE}
#Clear existing data and graphics
rm(list=ls())
graphics.off()

library(Hmisc)
library(dplyr)
library(tidyr)
```


```{r import-data, include=FALSE}
data=read.csv('2022BCMDatathonSurve_DATA_2022-11-03_1247.csv')
```


```{r labels-etc, echo=FALSE}
#Setting Labels

label(data$record_id)="Record ID"
label(data$redcap_survey_identifier)="Survey Identifier"
label(data$bcm_datathon_survey_2022_predatathon_knowledge_par_timestamp)="Survey Timestamp"
label(data$ar)="1. Academic rank"
label(data$ar_o)="Other, specify"
label(data$pdrc)="2. Primary department/section/research center"
label(data$yabcm)="3. Number of years at BCM"
label(data$phdaai)="1. I participated previously in a hackathon/datathon "
label(data$qipr)="2. I had experience using electronic medical record (EMR) data for research or quality improvement prior to this datathon "
label(data$pot)="1. Number of people on your datathon team"
label(data$wwdtm)="2. Have you worked with the members of your datathon team before"
label(data$role___1)="3. Your role on your team: (mark all that apply) (choice=Content (clinical) expert)"
label(data$role___2)="3. Your role on your team: (mark all that apply) (choice=Project lead)"
label(data$role___3)="3. Your role on your team: (mark all that apply) (choice=Chart reviewer)"
label(data$role___4)="3. Your role on your team: (mark all that apply) (choice=Data scientist)"
label(data$role___5)="3. Your role on your team: (mark all that apply) (choice=Statistical analyst)"
label(data$role___6)="3. Your role on your team: (mark all that apply) (choice=Data warehouse expert)"
label(data$role___7)="3. Your role on your team: (mark all that apply) (choice=Data manager)"
label(data$role___8)="3. Your role on your team: (mark all that apply) (choice=Learner)"
label(data$role___9)="3. Your role on your team: (mark all that apply) (choice=Other (please specify))"
label(data$orole)="Other, specify"
label(data$kadapr)="1. How knowledgeable were you about data availability at BCM prior to the datathon"
label(data$kadupr)="2. How knowledgeable were you about how to use a data warehouse prior to the datathon?"
label(data$kadlpr)="3. What was your understanding of a data warehouse limitations prior to the datathon"
label(data$bcm_datathon_survey_2022_predatathon_knowledge_par_complete)="Complete?"
label(data$bcm_datathon_survey_2022_postdatathon_knowledge_pa_timestamp)="Survey Timestamp"
label(data$hrdw)="1. Estimate how many person-hours you (personally) spent on datathon work, including time spent with BCM information technology (IT)"
label(data$prodw)="2. What percentage of the time spent on datathon work was with BCM IT"
label(data$doifda)="3. How difficult was obtaining information from a Data warehouse"
label(data$dwwda)="4. How difficult was working with a Data warehouse files"
label(data$kadaf)="1. How knowledgeable were you about data availability at BCM following the datathon"
label(data$kaduf)="2. How knowledgeable were you about how to use a data warehouse following the datathon"
label(data$kadlf)="3. What was your understanding of a data warehouse limitations following the datathon"
label(data$cpbeoda)="1. I completed my project by the end of the datathon"
label(data$sacq)="2. I successfully answered my clinical question"
label(data$cwiod)="3. I collaborated with individuals outside of my department"
label(data$cwni)="4. I collaborated with new individuals as part of the datathon"
label(data$pcfsuda)="5. I plan to conduct future studies using data from the BCM warehouse"
label(data$tpriasa)="6. Our team datathon project resulted in a scientific abstract for a local or national meeting"
label(data$tdprip)="7. Our team datathon project resulted in a publication"
label(data$pidaav)="8. Participating in the BCM datathon was a valuable experience"
label(data$wpfda)="9. I would participate in a future BCM datathon"
label(data$www)="10. What worked well or didnt work, or general comments"
label(data$bcm_datathon_survey_2022_postdatathon_knowledge_pa_complete)="Complete?"
#Setting Units


#Setting Factors(will create new variable for factors)
data$ar.factor = factor(data$ar,levels=c("1","2","3","4","5","6"))
data$phdaai.factor = factor(data$phdaai,levels=c("1","0"))
data$qipr.factor = factor(data$qipr,levels=c("1","2","3","4","5"))
data$wwdtm.factor = factor(data$wwdtm,levels=c("1","2","3"))
data$role___1.factor = factor(data$role___1,levels=c("0","1"))
data$role___2.factor = factor(data$role___2,levels=c("0","1"))
data$role___3.factor = factor(data$role___3,levels=c("0","1"))
data$role___4.factor = factor(data$role___4,levels=c("0","1"))
data$role___5.factor = factor(data$role___5,levels=c("0","1"))
data$role___6.factor = factor(data$role___6,levels=c("0","1"))
data$role___7.factor = factor(data$role___7,levels=c("0","1"))
data$role___8.factor = factor(data$role___8,levels=c("0","1"))
data$role___9.factor = factor(data$role___9,levels=c("0","1"))
data$kadapr.factor = factor(data$kadapr,levels=c("1","2","3","4","5"))
data$kadupr.factor = factor(data$kadupr,levels=c("1","2","3","4","5"))
data$kadlpr.factor = factor(data$kadlpr,levels=c("1","2","3","4","5"))
data$bcm_datathon_survey_2022_predatathon_knowledge_par_complete.factor = factor(data$bcm_datathon_survey_2022_predatathon_knowledge_par_complete,levels=c("0","1","2"))
data$doifda.factor = factor(data$doifda,levels=c("1","2","3","4","5"))
data$dwwda.factor = factor(data$dwwda,levels=c("1","2","3","4","5"))
data$kadaf.factor = factor(data$kadaf,levels=c("1","2","3","4","5"))
data$kaduf.factor = factor(data$kaduf,levels=c("1","2","3","4","5"))
data$kadlf.factor = factor(data$kadlf,levels=c("1","2","3","4","5"))
data$cpbeoda.factor = factor(data$cpbeoda,levels=c("1","2","3"))
data$sacq.factor = factor(data$sacq,levels=c("1","2","3"))
data$cwiod.factor = factor(data$cwiod,levels=c("1","0"))
data$cwni.factor = factor(data$cwni,levels=c("1","0"))
data$pcfsuda.factor = factor(data$pcfsuda,levels=c("1","2","3","4","5"))
data$tpriasa.factor = factor(data$tpriasa,levels=c("1","2","3"))
data$tdprip.factor = factor(data$tdprip,levels=c("1","2","3"))
data$pidaav.factor = factor(data$pidaav,levels=c("1","2","3","4","5"))
data$wpfda.factor = factor(data$wpfda,levels=c("1","2","3","4","5"))
data$bcm_datathon_survey_2022_postdatathon_knowledge_pa_complete.factor = factor(data$bcm_datathon_survey_2022_postdatathon_knowledge_pa_complete,levels=c("0","1","2"))

levels(data$ar.factor)=c("Student","Fellow","Assistant Professor","Associate Professor","Full Professor","Other (e.g. Staff, Instructor)")
levels(data$phdaai.factor)=c("Yes","No")
levels(data$qipr.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$wwdtm.factor)=c("Yes - all of the team member","Yes - some of the team members","No")
levels(data$role___1.factor)=c("Unchecked","Checked")
levels(data$role___2.factor)=c("Unchecked","Checked")
levels(data$role___3.factor)=c("Unchecked","Checked")
levels(data$role___4.factor)=c("Unchecked","Checked")
levels(data$role___5.factor)=c("Unchecked","Checked")
levels(data$role___6.factor)=c("Unchecked","Checked")
levels(data$role___7.factor)=c("Unchecked","Checked")
levels(data$role___8.factor)=c("Unchecked","Checked")
levels(data$role___9.factor)=c("Unchecked","Checked")
levels(data$kadapr.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$kadupr.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$kadlpr.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$bcm_datathon_survey_2022_predatathon_knowledge_par_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$doifda.factor)=c("1 (Easy)","2","3 (Moderate)","4","5 (Difficult)")
levels(data$dwwda.factor)=c("1 (Easy)","2","3 (Moderate)","4","5 (Difficult)")
levels(data$kadaf.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$kaduf.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$kadlf.factor)=c("1 (Very Little)","2","3 (Moderate)","4","5 (Great Deal)")
levels(data$cpbeoda.factor)=c("Yes","Partially","No")
levels(data$sacq.factor)=c("Yes","Partially","No")
levels(data$cwiod.factor)=c("Yes","No")
levels(data$cwni.factor)=c("Yes","No")
levels(data$pcfsuda.factor)=c("1 (Disagree)","2","3 (Neutral)","4","5 (Agree)")
levels(data$tpriasa.factor)=c("Yes","Planned","No")
levels(data$tdprip.factor)=c("Yes","Planned","No")
levels(data$pidaav.factor)=c("1 (Disagree)","2","3 (Neutral)","4","5 (Agree)")
levels(data$wpfda.factor)=c("1 (Disagree)","2","3 (Neutral)","4","5 (Agree)")
levels(data$bcm_datathon_survey_2022_postdatathon_knowledge_pa_complete.factor)=c("Incomplete","Unverified","Complete")
```


```{r renaming, include=FALSE}
data %>%
  rename(
    academic_rank = ar.factor,
    department = pdrc,  # not a factor
    years = yabcm,  # numeric
    prior_datathon = phdaai.factor,
    prior_ehr = qipr.factor,
    team_size = pot,  # numeric
    team_prior_work = wwdtm.factor,
    role_clinical = role___1.factor,
    role_lead = role___2.factor,
    role_reviewer = role___3.factor,
    role_datasci = role___4.factor,
    role_stats = role___5.factor,
    role_datawarehouse = role___6.factor,
    role_datamgr = role___7.factor,
    role_learner = role___8.factor,
    role_other = role___9.factor,
    
    # these have factor but NOT using because I want to graph.
    # 3 pre, 3 post, 6 succeed
    # Likerts are easy to understand the meaning of each number.
    pre_know_avail = kadapr,
    pre_know_use = kadupr,
    pre_know_limit = kadlpr,
    post_know_avail = kadaf,
    post_know_use = kaduf,
    post_know_limit = kadlf,
    
    # These are tricky because Redcap set up levels in non intuitive way.
    succeed_completed = cpbeoda,  # yes partially no
    succeed_answered = sacq,  # yes partially no
    succeed_collaborated_outside = cwiod,  # yes no
    succeed_collaborated_people = cwni, # yes no
    success_abstract = tpriasa,  # yes planned no
    success_paper = tdprip,  # yes planned no
    
    # now do with factors, easier to make tables
    f_completed = cpbeoda.factor,
    f_answered = sacq.factor,
    f_collab_out = cwiod.factor,
    f_collab_ppl = cwni.factor,
    f_abstract = tpriasa.factor,
    f_paper = tdprip.factor,

    # Back to ordinary variables where we WILL use factors.
    person_hours = hrdw,  # numeric
    percent_with_it = prodw,  # numeric
    difficult_obtain = doifda.factor,
    difficult_working = dwwda.factor,
    attitude_future_study = pcfsuda.factor,
    attitude_valuable = pidaav.factor,
    attitude_future_datathon = wpfda.factor
  ) %>%
  select (
    academic_rank,
    department,
    years,
    prior_datathon,
    prior_ehr,
    team_size,
    team_prior_work,
    
    role_clinical,
    role_lead,
    role_reviewer,
    role_datasci,
    role_stats,
    role_datawarehouse,
    role_datamgr,
    role_learner,
    role_other,
    
    pre_know_avail,
    pre_know_use,
    pre_know_limit,
    person_hours,
    percent_with_it,
    difficult_obtain,
    difficult_working,
    post_know_avail,
    post_know_use,
    post_know_limit,
    
    succeed_completed,
    succeed_answered,
    succeed_collaborated_outside,
    succeed_collaborated_people,
    success_abstract,
    success_paper,
    
    f_completed,
    f_answered,
    f_collab_out,
    f_collab_ppl,
    f_abstract,
    f_paper,
    
    attitude_future_study,
    attitude_valuable,
    attitude_future_datathon
  ) -> data_numeric
```


# Table 1 components

## Number of survey responses

```{r prelim}
dim(data_numeric)[1]
```

## Academic rank, Datathon experience, EHR experience

You can find the name of each variable that we are tabulating after the little
dollar sign.

```{r categorical}
table(data_numeric$academic_rank)
table(data_numeric$prior_datathon)
table(data_numeric$prior_ehr)


```


## Datathon roles

```{r categor-roles}
table(data_numeric$role_clinical)
table(data_numeric$role_lead)
table(data_numeric$role_reviewer)
table(data_numeric$role_datasci)
table(data_numeric$role_stats)
table(data_numeric$role_datawarehouse)
table(data_numeric$role_datamgr)
table(data_numeric$role_learner)
table(data_numeric$role_other)
```

## Percentiles of: years at BCM, team size, effort, IT effort

```{r continuous}
quantile(data_numeric$years)
quantile(data_numeric$team_size)
quantile(data_numeric$person_hours, na.rm = TRUE)
quantile(data_numeric$percent_with_it, na.rm = TRUE)
```


# Measures of team success

## Ignore (skip) This Figure.

Ignore this plot so far (skip ahead to tables), because I need to fix the way the
numeric coding maps to meaningful words. This is just an example of future look and
feel of the plot.

```{r stack-graph, echo=FALSE}
data_numeric %>%
  select(succeed_completed, succeed_answered, succeed_collaborated_outside, succeed_collaborated_people,
         success_abstract, success_paper) %>%
  rename(
    completed = succeed_completed,
    answered = succeed_answered,
    collab_outside = succeed_collaborated_outside,
    collab_ppl = succeed_collaborated_people,
    abstract = success_abstract,
    paper = success_paper
  ) %>%
  gather(`completed`, `answered`, `collab_outside`,
         `collab_ppl`, `abstract`, `paper`,
         key = "Dimension", value = "Success_nonfactor") %>%
  filter(!is.na(Success_nonfactor)) -> collab_success
  # mutate_at(vars(Success_nonfactor), ~ case_when(
  # 	. == 1 ~ "Fully",
  # 	. == 0.5 ~ "Partially",
  # 	. == 0 ~ "None")) %>%
  # mutate_at(vars(Dimension), ~ case_when(
  # 	. == "workedTeamReverse" ~ "Collab. totally\nnew team",
  # 	. == "collab.outside" ~ "Collaborated\noutside",
  # 	. == "collab.new" ~ "Collab. w/\nany new",
  # 	. == "completed" ~ "Completed\nproject",
  # 	. == "answered" ~ "Answered\nquestion",
  # 	. == "pub.abstract" ~ "Published\nabstract",
  # 	. == "pub.paper" ~ "Published\npaper",
  # 	TRUE ~ .)) %>%
  
  # mutate(Success = factor(Success_nonfactor, levels=c("Fully","Partially","None"), ordered=TRUE)) -> collab_success

ggplot(collab_success, aes(Dimension)) +
	geom_bar(aes(fill = as.factor(Success_nonfactor))) +
	labs(x="Datathon goal", title = "1 Yes, 2 Partly, 3 No", subtitle = 'Except when 1 yes, 2 no! What is 0?') +
	scale_fill_grey()

```

## Tables of self-reported success

Read this section rather than looking at the figure.

We tabulate the following questions in this order:

1. I completed my project by the end of the datathon
2. I successfully answered my clinical question
3. I collaborated with individuals outside of my department
4. I collaborated with new individuals as part of the datathon
5. Our team datathon project resulted in a scientific abstract for a local or national meeting
6. Our team datathon project resulted in a publication

Again, to confirm
which variable we're tabulating, look at the variable name after the dollar sign,
inside each grey rectangle of R code.



```{r tables-success}
table(data_numeric$f_completed)
table(data_numeric$f_answered)
table(data_numeric$f_collab_out)
table(data_numeric$f_collab_ppl)
table(data_numeric$f_abstract)
table(data_numeric$f_paper)
```



# Satisfaction/Attitudes

I plan to conduct future studies using data from the BCM warehouse"
label(data$tpriasa)="6. Our team datathon project resulted in a scientific abstract for a local or national meeting"
label(data$tdprip)="7. Our team datathon project resulted in a publication"
label(data$pidaav)="8. Participating in the BCM datathon was a valuable experience"
label(data$wpfda)="9. I would participate in a future BCM datathon"


```{r satisfaction}
table(data_numeric$attitude_valuable)
table(data_numeric$attitude_future_datathon)
table(data_numeric$attitude_future_study)


data_numeric %>%
  drop_na(attitude_valuable) %>%
  ggplot(aes(x=attitude_valuable)) +
  geom_bar() +
  labs(x="Level of agreement",
       title ="Participating in the datathon was a valuable experience")
data_numeric %>%
  drop_na(attitude_future_datathon) %>%
  ggplot(aes(x=attitude_future_datathon)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  labs(x="Level of agreement",
       title ="I would participate in a future datathon")
data_numeric %>%
  drop_na(attitude_future_study) %>%
  ggplot(aes(x=attitude_future_study)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  labs(x="Level of agreement",
       title ="I plan to conduct future studies using data from the BCM warehouse")
```



# Pre-post knowledge figure



 How knowledgeable were you about data availability at BCM prior to the datathon"


2. How knowledgeable were you about how to use a data warehouse prior to the datathon?"


3. What was your understanding of a data warehouse limitations prior to the datathon"



```{r simple-bar-pre-post}


#### USE

data_numeric %>%
  rename(Before = pre_know_use, After = post_know_use) %>%
  gather(`Before`, `After`, key="prepost", value="know.use") %>%
  drop_na(know.use) %>%
  ggplot(aes(x=know.use)) +
  geom_bar() +
  facet_grid(vars(prepost)) +
  labs(title="How knowledgeable were you about how to use a data warehouse?",
       x="Knowledge level")

##### AVAIL

data_numeric %>%
  rename(Before = pre_know_avail, After = post_know_avail) %>%
  gather(`Before`, `After`, key="prepost", value="know.use") %>%
  drop_na(know.use) %>%
  ggplot(aes(x=know.use)) +
  geom_bar() +
  facet_grid(vars(prepost)) +
  labs(title="How knowledgeable were you about data availability at BCM?",
       x="Knowledge level")


#### LIMIT

data_numeric %>%
  rename(Before = pre_know_limit, After = post_know_limit) %>%
  gather(`Before`, `After`, key="prepost", value="know.use") %>%
  drop_na(know.use) %>%
  ggplot(aes(x=know.use)) +
  geom_bar() +
  facet_grid(vars(prepost)) +
  labs(title="What was your understanding of data warehouse limitations?",
       x="Knowledge level")


```


Paired plot: To be completed

```{r pre-post-plot, eval=FALSE, include=FALSE}
jitter_absolute = 0.02
jx = jitter_absolute
jy = 5 * jitter_absolute


data_numeric %>%
gather(`pre_know_use`, `post_know_use`, key="prepost", value="know.use") %>%
mutate(numeric_prepost = case_when(prepost == "pre_know_use" ~ 0, prepost == "post_know_use" ~ 1),
	eps_x = runif(nrow(data_numeric) * 2, -1 * jx, jx),
	eps_y = runif(nrow(data_numeric) * 2, -1 * jy, jy)) -> use

# FIXME - need to get that "id" column working?

data_numeric %>%
gather(`pre_know_avail`, `post_know_avail`, key="prepost", value="know.avail") %>%
mutate(numeric_prepost = case_when(prepost == "pre_know_avail" ~ 0, prepost == "post_know_avail" ~ 1)) %>%
select(id, numeric_prepost, know.avail) -> available

data_numeric %>%
gather(`pre_know_limit`, `post_know_limit`, key="prepost", value="know.limit") %>%
mutate(numeric_prepost = case_when(prepost == "pre_know_limit" ~ 0, prepost == "post_know_limit" ~ 1)) %>%
select(id, numeric_prepost, know.limit) -> limit

use %>%
full_join(available, by=c("id", "numeric_prepost")) %>%
full_join(limit, by=c("id", "numeric_prepost")) -> gathered_all

gathered_all %>%
mutate(use = know.use + eps_y, avail = know.avail + eps_y, limit = know.limit + eps_y,
	nprepost = numeric_prepost + eps_x) %>%
select(id, nprepost, use, avail, limit) %>%
gather(`use`, `avail`, `limit`, key="dimension", value="knowledge") %>%
mutate_at(vars(dimension), ~ case_when(. == "use" ~ "How to use data warehouse",
	. == "avail" ~ "Data availability",
	. == "limit" ~ "Data warehouse limitations",
	TRUE ~ .)) -> double_gathered   # goes into one big facetted paired plot








## facet
ggplot(double_gathered, aes(x = nprepost, y = knowledge, group = id)) +
 geom_point() +
  geom_line() +
  scale_x_continuous(breaks = c(0,1), labels = c("Before", "After")) +
  labs(x = NULL, y = "Self-reported knowledge") +
  facet_wrap(vars(dimension), nrow = 2) -> facet
```

# Free text comments

```{r freetext}
data$www
```

